The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-08-01 paper/paper.Rnw
deleted a very slow line for a faster line
Implemented reardon approach - it's not very reliable

% pisa_listcol.Rdata is a dataset produced by the source() expression from the above
% code chunk
<<loading_data-recoding>>=
pisa_all <- read_rds("./data/pisa_listcol.Rdata")

@@

@@

# Function calculates the bottom 30th quantile for the bottom educated and the 70th quantile
# for the top educated. If the quantiles can't be estimated, it returns two NA's instead
quantile_missing <- function(df, weights, probs) {

    quan <- try(Hmisc::wtd.quantile(
      df$escs_trend,
      weights = df[[weights]],
      probs = probs
      ))

    if (any("try-error" %in% class(quan))) {
      return(c(NA, NA))
      } else {
     return(c(quan[1], quan[2]))
    }
}

@@

@@

      .x <-
        .x %>%
        dplyr::select(wave,
                      matches(paste0("^PV.*", test, "$")),
                      escs_dummy,
                      country,
                      one_of(weights_var),
                      AGE)

@@

      # Calculate median math score of all PV's
      .x$dv <- apply(.x[test_vars], 1, median, na.rm = T)
      .x$dv <- rowMedians(as.matrix(.x[test_vars]), na.rm = T)

@@

      mod2 <-
        lmer(adj_pvnum ~ escs_dummy + (1 + escs_dummy | country),
             data = .x,
             weights = .x[[weights_var]])

      # Take the country coefficients (absolute coefficients)
      country_coef <-
        coef(mod2)$country %>%
        rownames_to_column() %>%
        gather(escs_dummy, Mean, -rowname) %>%
        mutate(escs_dummy = dplyr::recode(escs_dummy,
                                          `(Intercept)` = "0",
                                          `escs_dummy` = "1"))

      # Take the absolute country standard errors
      se <-
        se.coef(mod2)$country %>%
        as.data.frame() %>%
        rownames_to_column() %>%
        gather(escs_dummy, s.e., -rowname) %>%
        mutate(escs_dummy = dplyr::recode(escs_dummy,
                                          `(Intercept)` = "0",
                                          `escs_dummy` = "1"))

      results <-
        inner_join(country_coef, se, by = c("rowname", "escs_dummy")) %>%
        rename(country = rowname) %>%
        arrange(country, escs_dummy)

      message(paste0(unique(.x$wave), " modeling done"))
      results
    })
}

@@

% Remember to uncomment when compiling sweave again, so the chunk remains intact.
<<modeling>>=
adapted_year_data <-
    map(pisa_all2$value, ~ {
      if (unique(.x$wave) == "pisa2000") {
        # pisa2000 has a different coding so here I recode 6 to 7 so that in all waves the top edu
        # is 7 and the bottom is 1
        .x <-
          mutate(.x, new_hisced = as.character(dplyr::recode(as.numeric(high_edu_broad), `6` = 7)))
      } else {
        .x <-
          mutate(.x, new_hisced = as.character(high_edu_broad))
      }
      .x
})

>>>>

% pisa_listcol.Rdata is a dataset produced by the source() expression from the above
% code chunk
<<loading_data-recoding>>=
pisa_all <- read_rds("./paper/data/pisa_listcol.Rdata")

@@

<<>>=

library(perccalc)

d <- map(1:6, ~ {

df <- pisa_all2$value[[.x]]

df_try <-
  df %>%
  filter(country == "France") %>%
  select(country, high_edu_broad, matches("^PV*.MATH$"), W_FSTUWT) %>%
  mutate(high_edu_broad = factor(high_edu_broad, ordered = TRUE))

math_score <-
  df_try %>%
  select_if(is_double) %>%
  as.matrix() %>%
  rowMedians(na.rm = TRUE)

df_try <-
  df_try %>%
  select(country, high_edu_broad, W_FSTUWT) %>%
  add_column(math_score)

perc_diff(df_try, high_edu_broad, math_score, percentiles = c(90, 10))
}) %>% 
  setNames(seq(2000, 2015, 3)) %>%
  enframe() %>%
  unnest() %>%
  slice(seq(1, 12, 2))

d %>%
  ggplot(aes(name, value)) +
  geom_point() +
  geom_line(aes(group = 1))

@@

@@

adapted_year_data <-
    map(pisa_all2$value, ~ {
      if (unique(.x$wave) == "pisa2000") {
        # pisa2000 has a different coding so here I recode 6 to 7 so that in all waves the top edu
        # is 7 and the bottom is 1
        .x <-
          mutate(.x, new_hisced = as.character(dplyr::recode(as.numeric(high_edu_broad), `6` = 7)))
      } else {
        .x <-
          mutate(.x, new_hisced = as.character(high_edu_broad))
      }
      .x
})

@@

      .x <-
        .x %>%
        dplyr::select(wave,
                      matches(paste0("^PV.*", test, "$")),
                      high_edu_broad,
                      country,
                      one_of(weights_var),
                      AGE) %>%
        mutate(high_edu_broad = factor(high_edu_broad, ordered = TRUE))

@@

      # Calculate median math score of all PV's
      .x$dv <- rowMedians(as.matrix(.x[test_vars]), na.rm = T)

@@

      error_perc_diff <- function(data) {
        result <- try(perc_diff(data, high_edu_broad, adj_pvnum, W_FSTUWT, percentiles = probs))

        if (class(result) == "try-error") {
          c(NA, NA)
        } else {
          result
        }
      }

      .x <- .x %>% filter(country %in% countries)

      mod2 <- 
        .x %>%
        split(.x$country) %>%
        map(~ error_perc_diff(.x)) %>%
        Filter(function(x) !all(is.na(x)), .)

      mod2 %>%
        reduce(bind_rows) %>%
        mutate(country = names(mod2))
    })
}

results_math <- test_diff(adapted_year_data, reliability_pisa, "MATH", c(0.1, 0.9))

results_math %>%
  setNames(years) %>%
  enframe(name = "year") %>%
  unnest() %>%
  filter(country != "United States") %>%
  mutate(low = difference - 1.96 * se,
         high = difference + 1.96 * se) %>%
  ggplot(aes(year, difference)) +
  geom_point() +
  geom_line(aes(group = country)) +
  geom_errorbar(aes(ymin = low, ymax = high)) +
  facet_wrap(~ country)

# .x <- adapted_year_data[[4]]
# .y <- reliability_pisa[3]
# test <- "MATH"
# probs <- c(0.9, 0.1)

@@

% Remember to uncomment when compiling sweave again, so the chunk remains intact.
<<modeling>>=
